Type-Safe Modular Parsing

### 2017-01-14 paper/sections/Inheritance.tex
References

Unfortunately, when parsing OO style ASTs, there are difficulties with
the two dimensions of extensibility. On the one hand it is naturally hard to
add new operations to the AST, as usual in the OO
style~\cite{wadler1998expression}. On the other hand
grammars are not modularly extensible, thus the addition of new
language constructs is problematic. Therefore,
it is perhaps not a surprise that parser generators that also generate
AST code (such as ANTLR~\cite{parr2013definitive}), normally generate code based on
visitors. Clearly this seems a better choice since generating code
based on the {\sc Interpreter} or {\sc Composite} pattern would
simply be too inflexible, due to the dual limitation on extensibility.

In the remainder of this section we show a technique for
defining \emph{extensible} parsing code. That is new
language constructs can be modularly added without changing
existing parsing code. The technique employs standard OO
style mechanisms such as subtyping and inheritance, as well as,
Packrat parsing combinators, and enables a better integration bettwen 
parsing and OO ASTs.

>>>>

Unfortunately, when parsing OO style ASTs, there are difficulties with
the two dimensions of extensibility. On the one hand it is naturally hard to
add new operations to the AST, as usual in the OO
style~\cite{wadler1998expression}. On the other hand
grammars are not modularly extensible, thus the addition of new
language constructs is problematic. Therefore,
it is perhaps not a surprise that parser generators that also generate
AST code (such as ANTLR~\cite{antlr1995}), normally generate code based on
visitors. Clearly this seems a better choice since generating code
based on the {\sc Interpreter} or {\sc Composite} pattern would
simply be too inflexible, due to the dual limitation on extensibility.

In the remainder of this section we show a technique for
defining \emph{extensible} parsing code. That is new
language constructs can be modularly added without changing
existing parsing code. The technique employs standard OO
style mechanisms such as subtyping and inheritance, as well as,
Packrat parsing combinators, and enables a better integration between
parsing and OO ASTs.
